State of the Practice for Lattice Boltzmann Method Software

### 2022-01-29 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Updates to section comparing our ranking to an estimate of the community's ranking

Answers \rqref{RQ_CompareHQ2Popular}. Table~\ref{repometrics} presents our LBM
software package rankings along with the repository ranking metrics of each
software package. Nine packages do not use GitHub, so they do not have a measure
of repository stars. Looking at the repository stars of the other 15 packages,
we can observe a slight pattern where packages that have been highly ranked by
our assessment do have more stars than lower ranked packages. The best ranked
package (ESPResSo) has the second most number of stars. The ninth ranked package
(Sailfish) has the most number of stars. Packages designated as lower quality
often do not use GitHub or GitLab, or have few stars. HemeLB only has 12 stars,
but its location recently moved from a different GitHub repository. Our
assessment of this package might not be accurate. The number of stars does not
necessarily represent the perceptions of the community, but for lack of an
alternative measure we will use it this way. Besides missing data, another
threat to the validity of this comparison is the varying ages of the
repositories. Older packages have been able to accumulate stars and watches for
longer than newer packages. The true quality of new packages may not be
reflected in their stars and watches.

The repository watches column contains even less data to compare since two of
the packages (Palabos, waLBerla) use GitLab, which does not include this metric.
The pattern that can be observed with this metric is very similar to that of the
stars metric. The ninth ranked package (Sailfish) has the most number of
watches. The best ranked package (ESPResSo) has the second most number of
watches.

@@

Table~\ref{rankingcomparison} compares the top 10 ranked LBM software packages
with a LBM package ranking made by a domain expert. Five (Palabos, OpenLB, LUMA,
Sailfish, waLBerla) of the top 10 ranked packages in this assessment are also
found in the domain expert's top 10 ranking. Interestingly, the first three of
these five packages are listed in the same order on both lists. Looking at the
remaining packages, the top two ranked packages in this state of the practice
assessment (ESPResSo, Ludwig) are not on the domain expert's list. Perhaps the
intended applications of these packages, complex fluids for Ludwig and soft
matter research for ESPResSo, do not align with the research interests of the
domain expert. This reasoning may also apply to eighth ranked Musubi. Moving
down the list, pyLBM and DL\_MESO(LBE) did not make the domain expert's top 10
list but were mentioned by the domain expert. Looking at the remainder of the
domain expert's list, ASL is a general purpose tool for solving partial
differential equations and may have fallen out of scope of the authoritative
lists that were used to identify the initial LBM software list. The domain
expert's fifth ranked package (ch4-project) was on the initial software list of
this assessment, but was removed due to a lack of documentation. Open FSI is a
new project that uses Palabos. It was made public the year before data was
collected for this state of the practice exercise. It may not have been on
authoritative lists due to its young age. Similarly, LIFE was also made public
recently. Finally, the domain expert's 10th ranked package (HemeLB) was ranked
21st in this state of the practice assessment.

\begin{table}
\begin{center}
	\begin{tabular}{ p{2cm}p{4.5cm}p{3.5cm}}
		\toprule
		Rank & This Assessment & Domain Expert\\
		\midrule
		1 & ESPResSo (N/A) & Palabos\\
		2 & Ludwig (N/A) & OpenLB\\
		3 & Palabos (1) & LUMA\\
		4 & OpenLB (2) & ASL\\
		5 & LUMA (3) & ch4-project\\
		6 & pyLBM (16) & Open FSI\\
		7 & DL\_MESO (LBE) (11) & WaLBerla\\
		8 & Musubi (N/A) & LIFE\\
		9 & Sailfish (9) & Sailfish\\
		10 & WaLBerla (7) & HemeLB\\		
		\bottomrule
	\end{tabular}
	\caption{Ranking Comparison} \label{rankingcomparison}
\end{center}
\end{table}

@@

>>>>

To address \rqref{RQ_CompareHQ2Popular}, Table~\ref{repometrics} compares our
LBM software package rankings against their popularity in the research
community, as estimated by repository stars and watches. Nine packages do not
use GitHub, so they do not have a measure of repository stars. Looking at the
repository stars of the other 15 packages, we can observe a pattern where
packages that have been highly ranked by our assessment tend to have more stars
than lower ranked packages. Our best ranked package (ESPResSo) has the second
most stars, while our ninth ranked package (Sailfish) has the highest number of
stars. The same correlation is observed in the repository watch column, although
this column contains less data, since two of the packages (Palabos, waLBerla)
use GitLab, which does not track watches. Consistent with the star measure, our
ninth ranked package (Sailfish) has the highest number of watches and our best
ranked package (ESPResSo) has the second most watches. Packages designated as
lower quality often do not use GitHub or GitLab, or have only a few stars/watches.
Although the details of the rankings differ, our measures show that following
best practices correlates with increased popularity in the scientific community.

@@

Although our ranking and the estimate of popularity are consistent, trying to
make more specific conclusions from this data would be problematic.  To start
with, using stars and watches may not accurately measure popularity or
necessarily represent the perceptions of the community. Even if the stars and
watches are a good proxy for popularity, they are an unfair measure since
repositories vary in age. Older packages have had more time to accumulate stars
and watches. For instance, HemeLB only has 12 stars, but it is young because (at
the time of measurement) HemeLB had recently moved from a different GitHub
repository. From the other side, our estimate of ranking by best practices is
also an approximation.  Besides our use of shallow measures, the assumption of
equal weighting between all qualities to create our overall ranking
(Section~\ref{Sec_OverallQuality}) does not represent reality.  For real
projects some qualities are going to be more important than others.

Although our ranking and the estimate of the community's ranking are not perfect
measures, they do suggest a correlation between best practices and popularity.
We do not know which comes first, the use of best practices or popularity, but
we do know that the top ranked packages share common practices.  The next
sections will explore how the practices of the LBM community compare to the
broader research software community.  We will also investigate the practices
from the top projects that others within the LBM community, and within the
broader research software community, can potentially adopt.

@@

\wss{Also compare to recommended practices for research software via the
checklists and other articles.}
